SAT Competition 2020\tnoteref{title}

### 2021-06-28 paper/main.tex
consolidations

In a few cases {\tt DRAT-trim} ran into the verification timeout of 40,000 seconds. 
In the Main track, only those unsatisfiable benchmark instances for which the proof produced by a solver 
could be validated at least by {\tt DRAT-trim} were considered solved by the solver. 
While there were several cases where {\tt cake\_lpr} ran out of resources, 
there was no case where {\tt DRAT-trim} would accept a proof and {\tt cake\_lpr} would not.

@@

\label{sec:computing}
The Main, No-Limits, and Planning tracks were run on the StarExec cluster \cite{starexec}
with computing nodes  equipped with Intel Xeon \SI{2.4}{\giga\hertz} processors 
and \SI{128}{\giga\byte} of memory.
The time limit enforced on each solver for solving an instance was 5,000 seconds. 
In the Main track, proof validation was limited to 40,000 seconds per proof.)
The solvers were allowed to use the full \SI{128}{\giga\byte} of RAM.\footnote{
Unfortunately, the memory limit of \SI{24}{\giga\byte}, that was used in the previous years,
was by mistake advertised on the competition web page prior to solver submission.
This could have resulted in some solvers not ``daring'' to use the full \SI{128}{\giga\byte}
in the competition. We do not, however, have concrete evidence to support this possibility.}

@@

To compile the set of 300 new instances, we first applied a hardness criterion by filtering out all instances solved by \solver{Minisat} in less than ten minutes.\footnote{We note that
the limit of ten minutes is again somewhat arbitrary.
This runtime hardness filter essentially aims to make sure that enough instances are included in the final benchmark set which allow for distinguishing in terms of relative performance between the best-performing competing solvers.
Unfortunately this limit was much greater that the requirements imposed for  ``interesting'' benchmark instances
by the BYOB rule. It could be more sensible to impose the same ten minutes limit also for an instance being ``interesting''. However, this would require more efforts
at least in terms of computation times from the solver authors in order to construct a set of interesting new benchmarks required by the BYOB rule.}
From the resulting 1,012 instances, in order to obtain a balanced benchmark set, 
we randomly selected $k$~instances per author using the value  $k$ which ensured that  the resulting set contains at least 300 instances.
This \emph{rule-based randomization} procedure is detailed as
Algorithm~\ref{algo:select}. Specifically,
we randomly selected seven satisfiable and seven unsatisfiable instances per author (Lines~3 and~4) and added instances of yet unknown result if this did not yield a total of 14~instances (Lines~5--7). 
Of the such obtained 308~instances, we randomly removed eight satisfiable instances, yielding a total of 114~satisfiable, 78~unsatisfiable and 108~instances of unknown satisfiability
status.

@@

A greedy set cover of the solved instances by solvers of the Main track is presented in Table~\ref{tab:gscMain}. 
In the first iteration, the solver which solved the highest number of instances is selected; in our case it was \solver{Kissat-sat} with 264 instances as we know already from Table~\ref{tab:results}.
With these 264 instances already covered,
\solver{CaDiCaL-alluip} is the best in further contributing to the set by additional $22$ instances in iteration $2$.
We can see that the further iterations tend to add very little,
with the final four iterations adding one instance each. Note that each solver
that managed to solve an instance uniquely (i.e., being the only solver that solved a particular instance)
shows up in the greedy set cover.
% for example the case of the last listed  
%\solver{exp-V-LGB-MLD-CBT-DL}.
% Highlighting solvers that uniquely solve an instance is a property of 
Indeed, the greedy set cover metric highlights solvers which are able to uniquely solve specific benchmark instances
and thereby contribute to the current state of the art.

@@

Interestingly, each of the first five tuples $T_{k \leq 5}$ contains exactly one of the three \solver{Kissat} variants. 
For $i < 5$ the relation $T_{i} \subset T_{i+1}$ holds only under projection to base solvers due to the fluctuating variants of \solver{Kissat}. 
The composition changes more strongly in $T_6$. 
Interestingly, we now have both variants $\{ \solver{Kissat-sat}, \solver{Kissat-unsat} \} \subset T_6$, and moreover it holds that $T_6 \subset T_7$.
All solvers in $T_{k \leq 7}$ are among the top-performing solvers which received awards in the Main track, with the only exception of $\solver{Scavel01} \in T_4 \cap T_5$ (cf.~Table~\ref{tab:results}).

@@

Contributions to the VBS can be captured by clustering the instance space into instance families.
We evaluate the runtimes of the three winning solvers of the Main track on those new families which are represented by at least $14$ instances (cf.~Table~\ref{tab:families})
and report their places and scores in Table~\ref{tab:eval-by-family}. 
Interestingly, the overall best solver \solver{Kissat-sat} is outperformed by the second and third ranked solvers \solver{Relaxed-newTech} and \solver{CMS-ccnr-lsids} on the \emph{Anti-Bandwidth}, \emph{Vlsat}, and \emph{Influence Maximization} families by a large margin.

\subsection{Similarity of Solvers}
To investigate the similarity of solvers from the Main track, we define a
similarity metric based on the measured runtimes. We start by removing $84$
benchmarks that have not been solved by any solver. For the rest, a PAR-2 score
is assigned to each instance for every solver,
% based on the \num{5000}-second timeout
\emph{i.e.}, we set a score of \num{10000} for unsolved instances.
Each solver $S$ is thus associated with the PAR-2 scores $S_{1},\dots,S_{316}$.
% This leaves us with $316$ floats
% $S_{1},\dots,S_{316}$ for each solver $S$.
The \emph{similarity} of two solvers $S$
and $S'$, normalized to the interval $[0,1]$ is defined
as:
% by Equation~\ref{eq-similarity}.

% \begin{equation}
% \label{eq-similarity}
\[
	similarity(S, S') =\quad 1 - \frac{\sum{|S_{i} - S'_{i}|}}{316 \cdot \num{10000}}
\]
% \end{equation}
We calculate the \emph{similarity} of the $30$ solvers with the
best average PAR-2 score in the Main track. The results are shown in
Figure~\ref{fig-similarity-main} as a \emph{heat map}, similar to the visualization in \cite{Balyo:2015:SATRace}. 
Additionally, the result
of hierarchically clustering the solvers based on their similarity is
illustrated as a \emph{dendrogram}. The height at which two solvers or clusters
are joined reflects how similar they are. For example, enabling \emph{trail saving}
in \solver{CaDiCaL-alluip} has no impact on the runtime, resulting in
a similarity above \num{0.999}. Therefore, the two solvers are joined low in the
dendrogram.
% A similar visualization was used for the results of the SAT Race 2015 \cite{Balyo:2015:SATRace}.

@@

To determine how similar an estimate is to the original ranking, we calculate
the \emph{Spearman's rank correlation coefficient} of the two rankings.
% and compute the correlation coefficient.
Spearman's rank correlation coefficient
for two rankings $r_{1}$ and $r_{2}$
is defined by the following equation:
% \begin{equation}
% \label{eq-spearman}
\[
    \rho = 1 - \frac{6\cdot\sum_{S}{(r_{1}(S) - r_{2}(S))^2}}{n(n^2-1)},
\]
% \end{equation}
where $n$ is the number of solvers in the Main track ($48$), and the rankings
$r_{\{1,2\}}(S)$ map a solver $S$ to its rank, \emph{i.e.,} $1$ for the best
performing solver, whereas $48$ is assigned to the solver
with the highest average PAR-2 score.

@@

The 2020 SAT Competition successfully continues the tradition of the SAT Competition series. In 2020, significant advances
in SAT solvers compared to previous years were observed. Some of the more interesting observations on the winning solving 
strategies include the following.
All winning solvers of the \emph{Main} track periodically schedule runs of a stochastic local search (SLS)
solver and import statistical information generated in unsuccessful SLS runs to reconfigure weights in their branching heuristics. As observed
from the results of the\emph{Parallel} track, it appears difficult to make proper use of more that $32$ threads for SAT solving, as in some occasions the $32$-threaded version of the same solver outperformed its $64$-threaded counterpart. 
However, from the winner in the massively parallel \emph{Cloud} track, we can learn that classical all-to-all clause-sharing can be outperformed by a more sophisticated clause-sharing architecture. 
It appears challenging to integrate and test sophisticated state-of-the-art methods in an incremental SAT solver and thus solvers usually disable parts of their features in the incremental use case. 
The winner of the \emph{Incremental Library} track shows that it is worth integrating a full solver functionality in the incremental use case.

@@

The first instantiation of an Application track was the Planning track organized as a one-time track in 2020. 
The results show that different solvers take the lead when we only evaluate for a single application, when compared to the overall Main track results.
We intend to run further instantiations of the Application track. 
While none of the solvers that participated in the Planning track seemed particularly optimized for planning instances, we hope that
in future iterations the community will pick up on the challenge of optimizing SAT solvers towards different focus applications, as a complementary
challenge when compared to the generality of the Main track.
In the next competition, the  planned focus of Application track will be on SAT solver applications in cryptology.

>>>>

In a few cases {\tt DRAT-trim} ran into the verification timeout of 45,000 seconds. 
In the Main track, only those unsatisfiable benchmark instances for which the proof produced by a solver 
could be validated at least by {\tt DRAT-trim} were considered solved by the solver. 
While there were several cases where {\tt cake\_lpr} ran out of resources, 
there was no case where {\tt DRAT-trim} would accept a proof and {\tt cake\_lpr} would not.

@@

\label{sec:computing}
The Main, No-Limits, and Planning tracks were run on the StarExec cluster \cite{starexec}
with computing nodes  equipped with Intel Xeon \SI{2.4}{\giga\hertz} processors 
and \SI{128}{\giga\byte} of memory.
The time limit enforced on each solver for solving an instance was 5,000 seconds. 
In the Main track, proof validation was limited to 45,000 seconds per proof.)
The solvers were allowed to use the full \SI{128}{\giga\byte} of RAM.\footnote{
Unfortunately, the memory limit of \SI{24}{\giga\byte}, that was used in the previous years,
was by mistake advertised on the competition web page prior to solver submission.
This could have resulted in some solvers not ``daring'' to use the full \SI{128}{\giga\byte}
in the competition. We do not, however, have concrete evidence to support this possibility.}

@@

To compile the set of 300 new instances, we first applied a hardness criterion by filtering out all instances solved by \solver{Minisat} in less than ten minutes.\footnote{Note that
the limit of ten minutes is again somewhat arbitrary.
This runtime hardness filter essentially aims to make sure that enough instances are included in the final benchmark set which allow for distinguishing in terms of relative performance between the best-performing competing solvers.
Unfortunately this limit was much greater that the requirements imposed for  ``interesting'' benchmark instances
by the BYOB rule. It could be more sensible to impose the same ten minutes limit also for an instance being ``interesting''. However, this would require more efforts
at least in terms of computation times from the solver authors in order to construct a set of interesting new benchmarks required by the BYOB rule.}
From the resulting 1,012 instances, in order to obtain a balanced benchmark set, 
we randomly selected $k$~instances per author using the value  $k$ which ensured that  the resulting set contains at least 300 instances.
This \emph{rule-based randomization} procedure is detailed as
Algorithm~\ref{algo:select}. Specifically,
we randomly selected seven satisfiable and seven unsatisfiable instances per author (Lines~3 and~4) and added instances of yet unknown result if this did not yield a total of 14~instances (Lines~5--7). 
Of the such obtained 308~instances, we randomly removed eight satisfiable instances, yielding a total of 114~satisfiable, 78~unsatisfiable and 108~instances of unknown satisfiability
status.

@@

A greedy set cover of the solved instances by solvers of the Main track is presented in Table~\ref{tab:gscMain}. 
In the first iteration, the solver which solved the highest number of instances is selected; in our case it was \solver{Kissat-sat} with 264 instances as we know already from Table~\ref{tab:results}.
With these 264 instances already covered,
\solver{CaDiCaL-alluip} is the best in further contributing to the set by additional $22$ instances in the second iteration.
We can see that the further iterations tend to add very little,
with the final four iterations adding one instance each. Note that each solver
that managed to solve an instance uniquely (i.e., being the only solver that solved a particular instance)
shows up in the greedy set cover.
% for example the case of the last listed  
%\solver{exp-V-LGB-MLD-CBT-DL}.
% Highlighting solvers that uniquely solve an instance is a property of 
Indeed, the greedy set cover metric highlights solvers which are able to uniquely solve specific benchmark instances
and thereby contribute to the current state of the art.

@@

Interestingly, each of the first five tuples $T_{k \leq 5}$ contains exactly one of the three \solver{Kissat} variants. 
The set $T_2$ is composed of the two winners of the Main SAT and Main UNSAT tracks. 
For $i < 5$ the relation $T_{i} \subset T_{i+1}$ holds only under projection to base solvers due to the fluctuating variants of \solver{Kissat}.

The composition changes more strongly in $T_6$. 
Interestingly, we now have both variants $\{ \solver{Kissat-sat}, \solver{Kissat-unsat} \} \subset T_6$, and moreover it holds that $T_6 \subset T_7$.
All solvers in $T_{k \leq 7}$ are among the top-performing solvers which received awards in the Main track, with the only exception of $\solver{Scavel01} \in T_4 \cap T_5$ (cf.~Table~\ref{tab:results}).

@@

Contributions to the VBS can be captured by clustering the instances by their family. 
We evaluate the runtimes of the three winning solvers of the Main track on those new families which are represented by at least $14$ instances (cf.~Table~\ref{tab:families})
and report their places and scores in Table~\ref{tab:eval-by-family}. 
Interestingly, the overall best solver \solver{Kissat-sat} is outperformed by the second and third ranked solvers \solver{Relaxed-newTech} and \solver{CMS-ccnr-lsids} on the \emph{Anti-Bandwidth}, \emph{Vlsat}, and \emph{Influence Maximization} families by a large margin.

\subsection{Similarity of Solvers}
To investigate the similarity of solvers from the Main track, we define a
similarity metric based on the measured runtimes. We start by removing $84$
benchmarks that have not been solved by any solver. For the rest, a PAR-2 score
is assigned to each instance for every solver,
% based on the \num{5000}-second timeout
i.e., we set a score of 10,000 for unsolved instances.
Each solver $S$ is thus associated with the PAR-2 scores $S_{1},\dots,S_{316}$.
% This leaves us with $316$ floats
% $S_{1},\dots,S_{316}$ for each solver $S$.
The \textsf{similarity} of two solvers $S$
and $S'$, normalized to the interval $[0,1]$ is defined
as:
% by Equation~\ref{eq-similarity}.

% \begin{equation}
% \label{eq-similarity}
\[
	\mathsf{similarity}(S, S') =\quad 1 - \frac{\sum{|S_{i} - S'_{i}|}}{316 \cdot \num{10000}}
\]
% \end{equation}
We calculate the \textsf{similarity} of the $30$ solvers with the
best average PAR-2 score in the Main track. The results are shown in
Figure~\ref{fig-similarity-main} as a \emph{heat map}, similar to the visualization in \cite{Balyo:2015:SATRace}. 
Additionally, the result
of hierarchically clustering the solvers based on their \textsf{similarity} is
illustrated as a \emph{dendrogram}. The height at which two solvers or clusters
are joined reflects how similar they are. For example, enabling \emph{trail saving}
in \solver{CaDiCaL-alluip} has no impact on the runtime, resulting in
a similarity above \num{0.999}. Therefore, the two solvers are joined low in the
dendrogram.
% A similar visualization was used for the results of the SAT Race 2015 \cite{Balyo:2015:SATRace}.

@@

To determine how similar an estimate is to the original ranking, we calculate
the \emph{Spearman's rank correlation coefficient} of the two rankings.
% and compute the correlation coefficient.
Spearman's rank correlation coefficient
for two rankings $r_{1}$ and $r_{2}$
is defined by the following equation:
% \begin{equation}
% \label{eq-spearman}
\[
    \rho = 1 - \frac{6\cdot\sum_{S}{(r_{1}(S) - r_{2}(S))^2}}{n(n^2-1)},
\]
% \end{equation}
where $n$ is the number of solvers in the Main track ($48$), and the rankings
$r_{\{1,2\}}(S)$ map a solver $S$ to its rank, i.e., $1$ for the best
performing solver, whereas $48$ is assigned to the solver
with the highest average PAR-2 score.

@@

The 2020 SAT Competition successfully continues the tradition of the SAT Competition series. In 2020, significant advances
in SAT solvers compared to previous years were observed. Some of the more interesting observations on the winning solving 
strategies include the following.
All winning solvers of the \emph{Main} track periodically schedule runs of a stochastic local search (SLS)
solver and import statistical information generated in unsuccessful SLS runs to reconfigure weights in their branching heuristics. As observed
from the results of the \emph{Parallel} track, it appears difficult to make proper use of more than $32$ threads for SAT solving, as in some occasions the $32$-threaded version of the same solver outperformed its $64$-threaded counterpart. 
However, from the winner in the massively parallel \emph{Cloud} track, we can learn that classical all-to-all clause-sharing can be outperformed by a more sophisticated clause-sharing architecture. 
It appears challenging to integrate and test sophisticated state-of-the-art methods in an incremental SAT solver and thus solvers usually disable parts of their features in the incremental use case. 
The winner of the \emph{Incremental Library} track shows that it is worth integrating a full solver functionality in the incremental use case.

@@

The first instantiation of an Application track was the Planning track organized as a one-time track in 2020. 
The results show that different solvers take the lead when we only evaluate for a single application, when compared to the overall Main track results.
We intend to run further instantiations of the Application track. 
While none of the solvers that participated in the Planning track seemed particularly optimized for planning instances, we hope that
in future iterations the community will pick up on the challenge of optimizing SAT solvers towards different focus applications, as a complementary
challenge when compared to the generality of the Main track.
In the next competition, the  planned focus of Application track will be on SAT solver applications in cryptography.
